---
title: "0_initialzie"
author: "Mehdi Salmani Jelodar - BCG"
date: "November 14, 2017"
output: html_document
---

# Churn Prediction Using Machine Learning 

Maximiliano Santinelli - santinelli.max@bcg.com
Mehdi Salmani Jelodar - salmani.mehdi@bcg.com


## Introduction

This tutorial builds models to predict churn using different methodologies - CART, rule-based models, C5.0, random forests (RF), artifitial neural networks (ANN), deep learning and ensemble of C5.0, RF and ANN models. A comparative assessment of the models performance is shown, based on their respective confusion matrix and ROC curves. 
The analyses is focused on the telco data set available in the [C50 R package](https://cran.r-project.org/web/packages/C50/index.html).



## Packages Installation

The first step is to install the packages required to chart the source and build the different predictive models.

```{r warning=FALSE, message=FALSE}
requiredPackages = c("ggplot2","corrplot","rpart","C50","randomForest","neuralnet","pROC")
#check the packages and if not installed install them and then load 
requiredPackages.check <- lapply(requiredPackages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})
#list all of the loaded packages
search()
```
```{r eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2) #graphics library
library(corrplot) #correlation charts
library(devtools) #required by bcgstyle
library(rpart) # CART model
library(C50) # C50 and Rule-Based classification models
library(randomForest) #Random forest model
library(neuralnet) #Neural network model
library(pROC) # ROC curve and AUC metric
```

## Data Load and Preparation

The data - a training and a test dataset, respectively - can be loaded directly from the C50 package into two separate data frames and briefly analyzed to understand the different data types. As you see we have 2 data sets: test and train. In machine learning (ML) applications, we usually split data into two subsets: a) training data set and b) testing data set. After training/fitting/preparing the ML model on the train data set, in order to make predictions/classificaiton/regression on the test data set to find out about underfitting or overfitting, which we need to avoid. 
#results='hide'}
```{r} 
# Load the test data available in the C50 package - create a test and a train data set
data(churn)

# Look at the shape of the data sets (count of rows and columns). 
dim(churnTrain)
dim(churnTest)

# Preliminary look at the Train data set
head(churnTrain)
summary(churnTrain)

# Look in detail at the target variable (churn)
head(churnTest$churn)

# Check the variable types of each column in Train and Test data sets. 
sapply(churnTrain, class)
sapply(churnTest, class)
```